---
title: "crc_ergothionase"
author: "Angela Jiang"
date: "`r Sys.Date()`"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(ggplot2)
library(dplyr)
library(sjPlot)
library(ggdist)
library(ggpubr)

```

## R Markdown

### Read in files

```{r read}
crc_mgx = read.csv("../Tables/diagnosis_crc_reads_all.tsv", sep="\t")
```

### Process data

Removing duplicates, remove samples with <1mil reads, normalizing by calculating count per million

```{r process}
crc_mgx_processed <- crc_mgx %>%
  filter(Total.Reads >= 1000000) %>%
  drop_na(diagnosis)



```

### Normality test
```{r normality, echo=FALSE}
shapiro.test(crc_mgx_processed$CPM)

```

### MGX plot

Plotting MGX data

```{r mgx_plot, echo=FALSE}
crc_mgx_processed$diagnosis <- factor(crc_mgx_processed$diagnosis, levels = c(
  "control",
  "adenoma",
  "CRC"
))

comparisons <- list(
  c("CRC", "control"),
  c("adenoma", "control"),
  c("CRC", "adenoma")
)

all_groups_studies <- crc_mgx_processed %>%
  filter(Study == "FengQ_2015" | Study == "Zeller_2014")

two_groups_studies <- crc_mgx_processed %>%
    filter(Study == "Vogtmann_2016" | Study == "YuJ_2015")

p <- ggplot(all_groups_studies, aes(x = diagnosis, y = CPM + 0.01, color = diagnosis)) +
  geom_jitter(width = 0.1, size = 1, alpha = 0.5, shape = 16) +
  geom_boxplot(width = 0.5, color = "black", alpha = 0.2, outlier.shape = NA, fill = NA) +
  scale_color_brewer(palette = "Dark2") +
  scale_y_log10() +
  xlab("Diagnosis") +
  ylab("CPM") +
  theme_classic() +
  facet_wrap(~ Study) +
  stat_compare_means(
    comparisons = comparisons,
    method = "wilcox.test",
    method.args = list(alternative = "greater"),
    label = "p.signifi",
    tip.length = 0.01,
size = 3 
  )

p2 <- ggplot(two_groups_studies, aes(x = diagnosis, y = CPM + 0.01, color = diagnosis)) +
  geom_jitter(width = 0.1, size = 1, alpha = 0.5, shape = 16) +
  geom_boxplot(width = 0.5, color = "black", alpha = 0.2, outlier.shape = NA, fill = NA) +
  scale_color_brewer(palette = "Dark2") +
  scale_y_log10() +
  xlab("Diagnosis") +
  ylab("CPM") +
  theme_classic() +
  facet_wrap(~ Study) +
  stat_compare_means(
    comparisons = list(c("CRC", "control")),  # ← this must be a list of length-2 vectors
    method = "wilcox.test",
    method.args = list(alternative = "greater"),
    label = "p.signifi",
    tip.length = 0.01,
    size = 3 )



p_all<- p / p2

p_all

ggsave(file="../../Plots/CRC_ergothionase_all.png", plot=p_all, width=6, height=7)


```

```
### Filter data into comparison groups

```{r mtx_stat, echo=FALSE}

#MGX
CRCvsHealthy_mgx <- crc_mgx_processed %>%
  filter(diagnosis == "CRC" | diagnosis == "healthy")

AdenomavsHealthy_mgx <- crc_mgx_processed %>%
  filter(diagnosis == "adenoma" | diagnosis == "healthy")

CRCvsAdenoma_mgx <- crc_mgx_processed %>%
  filter(diagnosis == "CRC" | diagnosis == "adenoma")
  
```

## MGX stats
### CRC vs healthy MGX wilcox test

```{r mgx_uc_nonibd, echo=FALSE}
wilcox.test(CPM~ diagnosis, 
            data = CRCvsHealthy_mgx,
            exact = FALSE, alternative="less")
```

### Adenoma vs healthy MGX wilcox test

```{r AdenomavsHealthy_mgx, echo=FALSE}
wilcox.test(CPM~ diagnosis, 
            data = AdenomavsHealthy_mgx,
            exact = FALSE, alternative="less")
```

### CRC vs UC Adenoma wilcox test

```{r CRCvsAdenoma_mgx, echo=FALSE}
wilcox.test(CPM~ diagnosis, 
            data = CRCvsAdenoma_mgx,
            exact = FALSE)
```

